Automatic Hadith Graphs using Arabic Morphological Analysis

### 2011-09-21 tex/hadith.ijcnlp.tex
some updates for code of bible genealogies

Bible contains numerous genealogical lists which are of interest to Bible scholars, historians and critics. 
The genealogies in the Bible trace key figures especially kings of the nation of Israel back to Adam; 
the genealogy of Jesus is also recorded in the New testament by two of his followers: Matthew and Luke.

The importance of such lists was profound for the old nation of Israel. They helped establish the kingship lineage, determining 
those eligible to work as priests and in the temple. The whole concept of land ownership was also tied to descent. Most importantly,
genealogical links were preserved due to the possible privilege of being the ancestor of the awaited Messiah.~\cite{catholicEncyclopedia:Online}

For many believers, genealogies are used as the main indicators of chronology especially when calculating the age of mankind~\cite{Thomas:92}.
They where always used to defend the historicity of the events and characters recorded in the Bible in contrast to those present in other religious 
texts of old nations which contain much fables and myths. However, it is known that when different genealogical lists overlap, they are
not always identical and contain some seeming contradictions. Those conflicts are in many cases due to nicknames and omissions and need to be examined
carefully and justified if possible.

%In many cases, such lists cover overlapping sections and when they do, they sometimes exhibit some seeming contradictions 
%due to nicknames and other reasons.
Detecting those conflicts is a task that has been performed by countless 
number of Bible students, theologians and historians~\cite{catholicEncyclopedia:Online,Belote:Online,completeBibleGenealogy:Online}. 
Their results have been published in articles and books; however, most of them are in written format 
and rarely display the genealogical graphs to make it easier for the reader to pinpoint those conflicts and visualize them.

In addition, being able to create a comprehensive genealogical graph from all biblical accounts greatly enhances its study,
since readers would be able to better understand a character's genealogical context and life span. Such comprehensive graphs can be tagged
with time durations between a father and his giving birth of his sons as is common in many biblical accounts. 
Attempts have been made for such endeavor, but resulted in manually performed unappealing pseudo-complete graphs as in \cite{Belote:Online,SoulLiberty:Online}. The main difficulty that hindered this task was the huge size of the resulting graphs whose layout becomes
challenging. This difficulty is also present in the hadith problem we undertook in this paper; however, we were able to tackle this issue by 
allowing for partial layout to be drawn on the fly as the user navigates in the POR which is already computed, 
instead of precomputed complete layouting of the graph on which plain scrolling is used to navigate. (check the accompanying software).
In all cases, it is noteworthy, that all genealogical trees/lineages available in any literature today are the product of manual work. 
In what follows, we explain how NaGEMA can be easily manipulated to be utilized to generate this information automatically.

Before presenting the necessary modifications for the state-machine used by NaGEMA, we present two representative biblical sections that
contain genealogical lists. These sections contain overlapping linages, however stated with different styles: \\

``This is the written account of {\bf Adam}'s line...

@@

When {\bf Enosh} had lived 90 years, he became the father of {\bf Kenan}. 
And after he became the father of {\bf Kenan}, {\bf Enosh} lived 815 years and had other sons and daughters. 
Altogether, {\bf Enosh} lived 905 years, and then he died.'' (Genesis 5:1-11) \\

``{\bf Adam}, {\bf Seth}, {\bf Enosh}, {\bf Kenan}, {\bf Mahalalel}, {\bf Jared}, {\bf Enoch}, 
{\bf Methuselah}, {\bf Lamech}, {\bf Noah}. 
The sons of {\bf Noah}: {\bf Shem}, {\bf Ham} and {\bf Japheth}.
The sons of {\bf Japheth}: {\bf Gomer}, {\bf Magog}, {\bf Madai}, {\bf Javan}, {\bf Tubal}, {\bf Meshech} and {\bf Tiras}.
The sons of {\bf Gomer}: {\bf Ashkenaz}, {\bf Riphath} and {\bf Togarmah}.
The sons of {\bf Javan}: {\bf Elishah}, {\bf Tarshish}, the {\bf Kittim} and the {\bf Rodanim}.
The sons of {\bf Ham}: {\bf Cush}, {\bf Mizraim}, {\bf Put} and {\bf Canaan}.
The sons of {\bf Cush}: {\bf Seba}, {\bf Havilah}, {\bf Sabta}, {\bf Raamah} and {\bf Sabteca}.
The sons of {\bf Raamah}: {\bf Sheba} and {\bf Dedan}. 
{\bf Cush} was the father of {\bf Nimrod}, who grew to be a mighty warrior on earth. ...'' (1 Chronicles 1:1-8) \\

@@

The problem at hand can be stated as follows: Given the whole text of the Bible detect all sections that have genealogical lists
and extract from them the names structured in a genealogical tree. The resemblance between this problem and that of hadith is evident; 
but still there are some differences that should be highlighted. 
First, in biblical genealogies, the individual names are simple in contrast to those present in hadith; 
consequently, the concept of a `name connector (NMC)' is unnecessary in the problem at hand.
Second, genealogies are not a flat sequence of narrators as in hadith but a branching tree of decedents; 
for this reason, it is necessary to detect not just the names in the text but also the node to which they are to be attached.
Those two differences are treated in the state-machine presented in Figure~\ref{f:bible} that contains all changes necessary to adapt NAGEMA to detect 
biblical genealogies.

@@

Emails have been one of the most important tool for communication in business, distant family members and friends.
In many cases, one email may trigger a thread of replies and forwards that may circulate between different people. It may even
end up after several replies being sent to someone new to the whole thread and he/she might consider navigating through
the lengthy list of quoted replies/forwards to get a better understanding of the conversations which have been going prior to his
receiving the email. However, the lengthy lists of conversations with some added indentation or symbols each possibly in a different
format may be a cumbersome task to do. That's why email clients such as gmail try organizing each conversation in a separate tab for
better use. Nonetheless, such clients are restricted to emails that are already available in the recipient's `Inbox' or `Sent Mails'.
In other words, even if the email client receives an email which has 
tens of separate replies/forwards but that are not previously available for the client, it 
fails to organize them and instead keeps them as one list of quoted conversations.

@@

(check if it can be applicable)

(Here is an example from wikipedia (\url{http://wikitravel.org/en/Silk_Road}):

-Northern Silk Route \\
Dunhuang to Hami \\
Hami to Turfan \\ 
Turfan to Korla \\
Korla to Kashgar

-Middle Silk Route \\ 
Dunhuang to Cherchen \\
Cherchen to Korla \\
Korla to Kashgar

-Southern (Jade) Silk Route \\
Dunhuang to Cherchen \\
Cherchen to Khotan \\
Khotan to Yarkand \\
Yarkand to Kashgar \\
)

In many cases, these itineraries contain some common places and merging them in a POR is much more illustrative than having them in text format.
Also, when electronic maps are available, such textual itineraries can automatically be turned into emphasized routes on maps.

@@

(briefly explain how this specific application has general value through contributing to the general task of entity extraction [learned names])

>>>>

Biblical genealogical lists describe family trees up to Adam and are of interest to Bible scholars including historians and critics. 
They trace key figures such as Israelite kings and prophets. Figure~\ref{f:gene1} shows an excerpt from Genesis 5:1-11.

\begin{figure}\centering
\resizebox{0.9\columnwidth}{!}{
\parbox[t]{1\columnwidth}{
\tiny
``This is the written account of {\bf Adam}'s line...

@@

When {\bf Enosh} had lived 90 years, he became the father of {\bf Kenan}. 
And after he became the father of {\bf Kenan}, {\bf Enosh} lived 815 years and had other sons and daughters. 
Altogether, {\bf Enosh} lived 905 years, and then he died.''
}}
\caption{Genealogical list from Genesis 5:1-11}
\end{figure}

\begin{figure}\centering
\resizebox{0.9\columnwidth}{!}{
\parbox[t]{1\columnwidth}{
\tiny
``{\bf Adam}, {\bf Seth}, {\bf Enosh}, {\bf Kenan}, {\bf Mahalalel}, {\bf Jared}, {\bf Enoch}, 
{\bf Methuselah}, {\bf Lamech}, {\bf Noah}. 
The sons of {\bf Noah}: {\bf Shem}, {\bf Ham} and {\bf Japheth}.
The sons of {\bf Japheth}: {\bf Gomer}, {\bf Magog}, {\bf Madai}, {\bf Javan}, {\bf Tubal}, {\bf Meshech} and {\bf Tiras}.
The sons of {\bf Gomer}: {\bf Ashkenaz}, {\bf Riphath} and {\bf Togarmah}.
The sons of {\bf Javan}: {\bf Elishah}, {\bf Tarshish}, the {\bf Kittim} and the {\bf Rodanim}.
The sons of {\bf Ham}: {\bf Cush}, {\bf Mizraim}, {\bf Put} and {\bf Canaan}.
The sons of {\bf Cush}: {\bf Seba}, {\bf Havilah}, {\bf Sabta}, {\bf Raamah} and {\bf Sabteca}.
The sons of {\bf Raamah}: {\bf Sheba} and {\bf Dedan}. 
{\bf Cush} was the father of {\bf Nimrod}, who grew to be a mighty warrior on earth. ...'' }}
\caption{Genealogical list from 1 Chronicles 1:1-8}
\end{figure}

%The importance of such lists was profound for the old nation of Israel. They 
The genealogical lists established the kingship lineage and determined
rights and privileges in the temple. Genealogical lists also determined descent with all its implications in biblical times
such as the concept of reverted ownership of sold real estate; every fifty years ownership of sold land reverted to the 
original owners and their inheritors. Genealogical lists were preserved due to the possible privilege of being the ancestor of the awaited Messiah.~\cite{catholicEncyclopedia:Online}

For many believers, genealogies are used as the main indicators of chronology especially when calculating the age of mankind~\cite{Thomas:92}.
Believers used them to defend the historicity of Biblical events and characters against other accounts from other 
texts thought of as fables and myths. 
However, separate genealogical lists that overlap suffer from consistency issues. 
For example, Matthew and Luke report two different genealogies of Jesus. 
Several biblical researchers studied the consistency of the lists~\cite{catholicEncyclopedia:Online,Belote:Online,completeBibleGenealogy:Online}, 
and published articles and books on the topic. Up to our knowledge very few include graphical illustrations such as family trees and graphs 
which makes it easier for the reader to visualize the differences. 
Our methodology can automatically create such illustrations.

create a comprehensive genealogical graph from all biblical accounts greatly enhances its study,
since readers would be able to better understand a character's genealogical context and life span. Such comprehensive graphs can be tagged
with time durations between a father and his giving birth of his sons as is common in many biblical accounts.

Work such as~\cite{Belote:Online,SoulLiberty:Online} constructed partial genealogical graphs manually from several lists.
This work is short from being comprehensive.

In what follows, we explain how NaGEMA can be easily manipulated to be utilized 
to generate this information automatically.

%The main difficulty that hindered this task was the huge size of the resulting graphs whose layout becomes
%challenging. This difficulty is also present in the hadith problem we undertook in this paper; however, we were able to tackle this issue by 
%allowing for partial layout to be drawn on the fly as the user navigates in the POR which is already computed, 
%instead of precomputed complete layouting of the graph on which plain scrolling is used to navigate. (check the accompanying software).
%In all cases, it is noteworthy, that all genealogical trees/lineages available in any literature today are the product of manual work.

%Before presenting the necessary modifications for the state-machine used by NaGEMA, we present two representative biblical sections that
%contain genealogical lists. 
%NaGEMA can handle 
%These sections contain overlapping linages, however stated with different styles: \\

@@

Given the whole text of the Bible, NaGEMA detects all sections that have genealogical lists
and extract from them the names structured in a genealogical DAG. The resemblance between this problem and that of hadith is evident; 
but still there are some differences that should be highlighted. 
First, in biblical genealogies, the individual names are simple in contrast to those present in hadith; 
consequently, the concept of a `name connector (NMC)' is unnecessary in the problem at hand.
Second, genealogies are not a flat sequence of narrators as in hadith but a branching tree of decedents; 
for this reason, it is necessary to detect not just the names in the text but also the node to which they are to be attached.
Those two differences are treated in the state-machine presented in Figure~\ref{f:bible} that contains all changes necessary to adapt NAGEMA to detect 
biblical genealogies.

@@

Emails have been one of the most important tool for communication in business, distant family members and friends.
In many cases, one email may trigger a thread of replies and forwards that may circulate between different people. It may even
end up after several replies being sent to someone new to the whole thread and he/she might consider navigating through
the lengthy list of quoted replies/forwards to get a better understanding of the conversations which have been going prior to 
receiving the email. However, the lengthy lists of conversations with some added indentation or symbols each possibly in a different
format may be a cumbersome task to do. That's why email clients such as Gmail try organizing each conversation in a separate tab for
better use. Nonetheless, such clients are restricted to emails that are already available in the recipient's `Inbox' or `Sent Mails'.
In other words, even if the email client receives an email which has 
tens of separate replies/forwards but that are not previously available for the client, it 
fails to organize them and instead keeps them as one list of quoted conversations.

@@

Travel Itineraries are plans of travel or routes. They are usually used as guides for travelers. They list different
places through which the route to the destination passes. 
In many cases, those guides are provided in written format
It is also common that alternative routes be suggested to the same destination.
Consider the example below for the itineraries known as the `Silk Road'\footnote{from \url{http://wikitravel.org/en/Silk_Road}}:

{\setlength{\parindent}{20pt} 
\setlength{\parskip}{1ex}
\indent-Northern Silk Route \\
\indent \indent Dunhuang to Hami \\
\indent \indent Hami to Turfan \\ 
\indent \indent Turfan to Korla \\
\indent \indent Korla to Kashgar

\indent -Middle Silk Route \\ 
\indent \indent Dunhuang to Cherchen \\
\indent \indent Cherchen to Korla \\
\indent \indent Korla to Kashgar

\indent -Southern (Jade) Silk Route \\
\indent \indent Dunhuang to Cherchen \\
\indent \indent Cherchen to Khotan \\
\indent \indent Khotan to Yarkand \\
\indent \indent Yarkand to Kashgar \\ }

In a similar manner to extracting hadith, a controller that looks for aggregations of Place names
with the NRC equivalent being phrases such as `to', `pass in', `stop at', etc... 
Afterwards, itineraries can be merged into a POR by merging common locations.
Clearly, such a display is much more illustrative than having them in text format.
Also, when electronic maps are available, POR's can automatically be turned into emphasized routes on maps.

@@

The problem at hand has significant contribution to the general task of Named Entity Recognition (NER). It has been noted that
gazetteers play an important role in the accuracy of knowledge-based approaches and bootstrapping learning
approaches~\cite{Kozareva:06,Carlson:09,Toral:06,Benajiba:07,Kazama:08,Nadeau:06}. Nonetheless, although gazetteers result in high accuracy,
they are hard to obtain and mantain. % They are also language and domain-dependent. 
Consequently, many automatic methods have been proposed to generate extensive gazetteers instead of relying on the time-consuming manual work
to compile such lists. Proposed methods include: writing some rules for the context of the named entities targeted such as propositions that
introduce them, also known as trigger gazetteers~\cite{Kozareva:06,Toral:06}; utilizing free online encyclopedias such as Wikipedia which are 
continually maintained~\cite{Benajiba:07,Kazama:08}; in addition to reliance on web search engines starting from a small list of words and extracting
similar that appear in the extracted pages~\cite{Nadeau:06}.

As previously noted, when solving the hadith problem, we were able to learn much previously unknown names using their context. We have used methods
relating to those found in~\cite{Traboulsi:09,Kozareva:06}; however, our learned lists appear to be more accurate than their counterparts due to the
the more characteristics of the hadith corpus and the wider context on which we rely on. Particularly, we learn only names that appear in the hadith sanad
detected through the presence of an aggregation of previously known names in its neighborhood along with name connectors such as narration verbs and 
prepositions. In Table X, we compare the results of our learning methods to evaluate the role of the sand context. Clearly, the sanad context provides 
a higher confidence in the accuracy of our extracted names. Accordingly, our generated gazetteers are superior than those extracted otherwise and 
enrich the task of Arabic Named Entity Extraction resources. The same applies to that of Place names that can be discovered in the context of travel
itineraries and the people names that appear in biblical genealogical lists. Since the latter sources are available in multitude of languages, they
can be a rich source for building gazetteers for resource-poor languages.

(present the necessary table and expand upon its significance)
